---
title: "Autocorrelation in R"
author: "Bidyut Kr. Ghosh"
date: "13 August 2017"
output: slidy_presentation
---

```{r setup, include=TRUE, comment=NA}
knitr::opts_chunk$set(echo = FALSE, comment = NA)
```

<style>
body {
text-align: justify}
</style>

# Data Description and Model Estimation

We will consider the Blaisdell Company example from page 489 of Applied Linear Regression Models (4th ed) by Kutner, Nachtsheim, and Neter.  
We will try to explain the company sales, ‘comsales’ ($ millions) with the help of single predictor variable, viz. ‘indsales’ by fitting a simple linear regression model.  
The data is given in the text format (blaisdell.txt).  
Mathematically we write the model as follows:  
 $$ {comsales}_t = \beta_0 +\beta_1 ({indsales}_t) +u_t$$ 
- Importing the data into R

```{r,include=TRUE}
blaisdell<-read.table(file = "/media/bidyut/HP-G6/RESEARCH/RESEARCH METHODOLOGY/Book_Publication/DATA SETS/blaisdell.txt" ,header = TRUE)
# To view the dataset
View(blaisdell)
# To attach the dataset into R-session
attach(blaisdell)
```

- To fit the initial model to explain company sales with industry sales

```{r,include=TRUE}
fit1<-lm(comsales~indsales)
summary(fit1)
```

# Residual Diagnostic

- To save the residuals(uhat1) of the model 

```{r,include=TRUE}
blaisdell$uhat1<-resid(fit1)
attach(blaisdell)
```

- To plot the residuals against time

```{r,include=TRUE}
plot(uhat1, type = "l")
library(ggplot2)
ggplot(blaisdell, aes(time, uhat1))+geom_line()
```


# Durbin-Watson test

```{r,include=TRUE}
library(car)
durbinWatsonTest(fit1)
```


# Remedial Measures: Cochrane-Orcutt Procedure 

```{r,include=TRUE}
library(orcutt)
fit2<-cochrane.orcutt(fit1)
summary(fit2)
```

# Remedial Measures: first-difference transformation model with no intercept term

```{r,include=TRUE}
fit3<-lm(diff(comsales,1)~0+diff(indsales,1))
summary(fit3)
```

-Durbin-Watson test on the first-difference model

```{r,include=TRUE}
durbinWatsonTest(fit3)
```

# Advanced Autocorrelation Problem using the dataset "us_consumption.xls"

- Importing the dataset using "readxl" package

```{r,include=TRUE}
library(readxl)
us_consumption<-read_excel("/media/bidyut/HP-G6/RESEARCH/RESEARCH METHODOLOGY/Book_Publication/DATA SETS/us_consumption.xls" , sheet = "Sheet1", col_names = TRUE,col_types = NULL, skip = 0)
# To view the dataset
View(us_consumption)
# To attach the dataset into Rsession
attach(us_consumption)
```


# Now we run the model to explain logarithm of consumption using logarithms of X variables in R and the results are shown below:

```{r,include=TRUE}
fit4<- lm(lnconsump~lndpi+lnwealth+interest)
summary(fit4)
```


# To save the residuals of the model fit

```{r,include=TRUE}
us_consumption$uhat1<-ts(resid(fit4))
attach(us_consumption)
```


# To plot the residuals of the model againt time

```{r,include=TRUE}
plot(uhat1, type="l")
library(ggplot2)
ggplot(us_consumption, aes(time, uhat1))+geom_line()
```


# To create the one-period lag of the saved residuals

```{r,include=TRUE}
us_consumption$luhat1<-lag(uhat1, -1)
```


# To plot the Current vs Lagged Residuals using basic graphics

```{r,include=TRUE}
attach(us_consumption)
plot(uhat1[1:54], luhat1[2:55], pch="*", main = "Current vs. Lagged Residuals")
abline(lm(luhat1[2:55]~uhat1[1:54]), col="red")
```


# To plot the Current vs Lagged Residuals using advanced graphics

```{r,include=TRUE}
ggplot(us_consumption, aes(luhat1[2:55], uhat1[1:54]))+geom_point()+geom_smooth(method = "lm", se=FALSE)
```



# To conduct the Durbin-Watson test

```{r,include=TRUE}
library(car)
durbinWatsonTest(fit4)
```


# To perform the Breusch–Godfrey (BG) test of autocorrelation.

```{r,include=TRUE}
library(lmtest)
bgtest(fit4, order = 1, type = "Chisq")
```


# Remedial Measures for Autocorrelation: Using the first difference method

```{r,include=TRUE}
fit5<- lm(diff(lnconsump,1)~0+diff(lndpi,1)+diff(lnwealth,1)+diff(interest,1))
summary(fit5)
```



# Breusch–Godfrey (BG) test of autocorrelation for the first difference model

```{r,include=TRUE}
bgtest(fit5,order = 1, type = "Chisq")
```

# Remedial Measures for Autocorrelation: Generalized Transformation

- To save the residuals of the initial model (fit4)

```{r,include=TRUE}
uhat4<-resid(fit4)
```

- To generate the one-period lag of the residuals.

```{r,include=TRUE}
library(dplyr)
us_consumption$luhat4<-lag(uhat4, n=1L)
```


# For estimating the rho

```{r,include=TRUE}
attach(us_consumption)
fit6<-lm(uhat4~0+luhat4)
summary(fit6)
```


# Thus, rho hat is found to be 0.3247. Further, we can recheck it approximately using D-W statistic

```{r,include=TRUE}
durbinWatsonTest(fit4)
```


# Complete transformation

```{r,include=TRUE}
us_consumption$Cstar<- lnconsump-0.3247*lag(lnconsump,n=1L)
us_consumption$Ystar<-lndpi-0.3247*lag(lndpi,n=1L)
us_consumption$Wstar<-lnwealth-0.3247*lag(lnwealth,n=1L)
us_consumption$rstar<-interest-0.3247*lag(interest,n=1L)
attach(us_consumption)
```

# To estimate the transformed model  

```{r,include=TRUE}
fit7<-lm(Cstar~Ystar+Wstar+rstar)
summary(fit7)
```


